The Journey from Wild to Text Book Data: A Case Study of the National Longitudinal Survey of Youth

### 2021-04-07 report_alt.Rmd
Add abstract.

>>>>

The NSLY79 is a prominent open data source that has been playing an important role in multidisciplinary research. This article aims to demonstrate the steps of initial data analysis (IDA), namely tidying, cleaning, and documenting the process to make the data available for text book or research data. Besides, this article provides the opportunity to continuously refresh the text book data by updating the codes that are provided in this article. The IDA shows that the NLSY79 data has two problems. Firstly, the downloaded data format is untidy. Secondly, some anomalies in terms of extreme high hourly wages are found. We addressed the first problem by tidying the data according to three principles of a tidy data. For the latter problem, we found the anomalies using the robustness weight threshold of robust linear regression. An observation with weight < 1.12 is considered as an outlier. Further, we fixed the anomalies using the predicted value of the robust linear regression model. Using the cleaned data, we provided three data sets, the high school mean hourly wage data, the high school dropouts mean hourly wage data, and demographic variables of the NLSY79 cohort. These data sets are contained in an open source R package `yowie`. Supplementary materials for this article are available online.

**Keywords**: Data cleaning; Data tidying; Longitudinal data; NLSY79; Open data; Initial data analysis; Outlier detection; Robust linear regression
